% !TeX root = main.tex
\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{fancyhdr}       % header
\usepackage{graphicx}       % graphics
\usepackage{bookmark}
\usepackage{enumitem}

%\usepackage{caption}
\usepackage{svg}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{flushend}
\usepackage{mathtools}
\usepackage[square,sort,comma,numbers]{natbib}
\bibliographystyle{IEEEtranN}
\usepackage{tabularray}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}
\usepackage{subfig}
\usepackage{listings}
% \usepackage{todonotes}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{pdfpages}
\usepackage{placeins}
\usepackage{cellspace}
\usepackage{todonotes}
\usepackage{mdframed}
\usepackage{makecell}
\usepackage[most]{tcolorbox}

\definecolor{lightyellow}{rgb}{1.0, 0.95, 0.8}

\newcommand{\Highlight}[1]{%
  \colorbox{lightyellow}{\parbox{0.8\linewidth}{#1}}}

\lstdefinestyle{javacode}{
    language=Java,
    basicstyle=\scriptsize\ttfamily,
    numbers=none,
    escapeinside={(*@}{@*)}
}
\newtcolorbox{findingbox}[1][]{
  colback=gray!10!white,
  colframe=gray!50!black,
  #1
}


\setlength\cellspacetoplimit{3.5pt}
\setlength\cellspacebottomlimit{3.5pt}

% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
%     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

	\title {Detecting Exception-Related Behavioural Breaking Changes with UnCheckGuard}

    \author{\IEEEauthorblockN{Vinayak Sharma\IEEEauthorrefmark{1}, Patrick Lam\IEEEauthorrefmark{2}}
    \IEEEauthorblockA{\IEEEauthorrefmark{1}University of Waterloo; \emph{vinayak.sharma1@uwaterloo.ca}}
    \IEEEauthorblockA{\IEEEauthorrefmark{2}University of Waterloo; \emph{patrick.lam@uwaterloo.ca}}
    }

    \maketitle
    \thispagestyle{plain}
    \pagestyle{plain}

    \begin{abstract}
      The ubiquitous use of third-party libraries in software development has enabled developers to quickly add
      new functionality to their client software. Unfortunately, library usage also carries a cost in
      terms of software maintenance: library upgrades may include breaking changes, in which client expectations
      about library behaviour are no longer met in new library versions. Behavioural breaking
      changes can be particularly insidious, and in their full generality, could require sophisticated program
      analysis techniques to (approximately) detect.

      In this work, we present our UnCheckGuard tool, which detects a class of behavioural breaking changes---those
      related to exceptions thrown by Java libraries. UnCheckGuard analyzes both sides of the library/client
      duet. On the library side, UnCheckGuard creates a list of new exceptions that may be thrown by methods
      in a library's public API, including by its transitive callees. On the client side, UnCheckGuard identifies
      client methods that call library methods with new exceptions. To reduce false positives, UnCheckGuard
      additionally filters out new exceptions that cannot be triggered by particular clients, using taint analysis. It therefore can be
      used by client developers as a tool to screen library updates for relevant incompatibilities.

      We have evaluated UnCheckGuard on 302 libraries and 352 library-client pairs drawn from the DUETS collection
      and found 120 libraries with newly-added exceptions, as well as 1708 callsites to library methods which,
      when upgraded to the latest version, may introduce
      a behavioural breaking change in the client due to a newly added unchecked exception. These findings
      highlight the practical value of UnCheckGuard in identifying exception-related incompatibilities
      introduced by library upgrades.


    \end{abstract}
    
    

    \begin{IEEEkeywords}
        client/library interactions, behavioural breaking changes, exceptions, static analysis
    \end{IEEEkeywords}


    \section{Introduction} \label{sec:introduction}
    \input{01-introduction}

    \section{Background}\label{sec:background}
    \input{02-background}


    \section{Motivating Example} \label{sec:motivating}
    \input{03-motivating-example}

    \section{Data Collection}\label{sec:data-collection}
    \input{04-data-collection}


    \section{Methodology}\label{sec:methodology}
    \input{05-methodology}


    \section{Results}\label{sec:results}
    \input{06-results}


    % TODO: add discussion
    % TODO: add threats to validity

    %\section{Threats to Validity}\label{sec:threats}
    %\input{threats}
    
    \section{Related Work}\label{sec:related-work}
    \input{07-related-work}
    

    \section{Conclusion}\label{sec:conclusion-&-future-work}
    \input{08-conclusion}
    
    \bibliography{references}

\end{document}
